---
title: "William Kumler"
knit: (function(inputFile, encoding) {
    rmarkdown::render(input = inputFile,
      encoding = encoding,
      output_format = c("word_document", "github_document", "vitae::awesomecv"),
      output_file = c("william_kumler.docx", "README.md", "william_kumler.pdf")
    )
  })
name: William
surname: Kumler
position: "Ph.D. Candidate"
address: "Seattle, WA"
email: "wkumler@uw.edu"
github: wkumler
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
word <- opts_knit$get("rmarkdown.pandoc.to")=="docx"
md <- opts_knit$get("rmarkdown.pandoc.to")=="gfm+tex_math_dollars-yaml_metadata_block"
latex <- opts_knit$get("rmarkdown.pandoc.to")=="latex"

eleg_space <- function(){
  word <- opts_knit$get("rmarkdown.pandoc.to")=="docx"
  md <- opts_knit$get("rmarkdown.pandoc.to")=="gfm+tex_math_dollars-yaml_metadata_block"
  # No idea why ifelse works when if;else doesn't but ok
  ifelse(md|word, ifelse(word, " \n<br>", " \n"), "\\hfill")
}
```

```{r, results='asis', include=FALSE}
alignment <- ifelse(word, "c", "l")
if(word|md){
  print(kable(data.frame(location="Seattle, WA"), col.names=NULL, align = alignment))
  metadata <- data.frame(
    email="wkumler@uw.edu",
    github="[github.com/wkumler](github.com/wkumler)",
    orcid="[ORCiD](https://orcid.org/0000-0002-5022-8009)"
  )
  print(kable(metadata, col.names = NULL, align=alignment))
}
```

```{r, results='asis'}
if(word|md){
  cat(paste(
    "Email: wkumler@uw.edu",
    "GitHub: [wkumler](https://github.com/wkumler)",
    "ORCiD: [0000-0002-5022-8009](https://orcid.org/0000-0002-5022-8009)",
    sep = "\n\n"
  ))
}
```

---

## Professional Summary

Ph.D. candidate in chemical oceanography focusing on marine microbial metabolomics under Dr. Anitra Ingalls. Expertise in automated and untargeted liquid chromatography mass spectrometry, with significant experience developing computational tools for mass spectrometry data analysis.

---

## Education

**Ph.D. in Chemical Oceanography**, University of Washington `r eleg_space()` _2019 - 2025_

**M.Sc. in Oceanography**, University of Washington `r eleg_space()` _2019 - 2023_

**B.A. in Marine Science**, University of California, Berkeley `r eleg_space()` _2014 - 2018_

**B.S. in Molecular and Environmental Biology**, University of California, Berkeley `r eleg_space()` _2014 - 2018_

---

## Work Experience

**Ph.D. student**, Ingalls Lab, University of Washington, Seattle `r eleg_space()` _2019 - 2025_

- Published papers on marine microbial metabolomics with a focus on untargeted methodologies
- Developed existing and novel mass spectrometry tools for analysis and visualization
- Fieldwork in the North Pacific Subtropical Gyre and California Current
- Mentored graduate and undergraduate students in metabolomics

**Lab Manager**, Koehl Lab, University of California, Berkeley `r eleg_space()` _2018 - 2019_

- Maintained live cultures of choanoflagellates and protozoa
- Managed and organized lab members, materials, and safety protocols
- Assisted visiting researchers with statistical analyses
- Tutored undergraduates in ImageJ and R programming

**Lab Manager**, Edwards Lab, University of California, Berkeley `r eleg_space()` _2018 - 2019_

- Performed HPLC-MS sample preparation and analysis using Thermo Fusion Lumos ID-X
- Developed and maintained a lipidomics pipeline for data handling and statistical analysis
- Educated undergraduates in R programming and SLURM cluster usage

---

## Publications

- **W. Kumler**, W. Qin, R.A. Lundeen, B. Barone, L.T. Carlson & A.E. Ingalls (2024). Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre. _Frontiers in Marine Science_, 11:1481409. [DOI:10.3389/fmars.2024.1481409](https://doi.org/10.3389/fmars.2024.1481409)
- **Kumler, W.**, Hazelton, B.J. & Ingalls, A.E. (2023). Picky with peakpicking: assessing chromatographic peak quality with simple metrics in metabolomics. _BMC Bioinformatics_, 24:404. [DOI:10.1186/s12859-023-05533-4](https://doi.org/10.1186/s12859-023-05533-4)
- **Kumler, W.** & Ingalls, A.E. (2022). Tidy data neatly resolves mass-spectrometry's ragged arrays. _R Journal_, 14(3), 193-202. [DOI:10.32614/RJ-2022-050](https://doi.org/10.32614/RJ-2022-050)
- Boysen, A.K., Durham, B.P., **Kumler, W.**, et al. (2022). Glycine betaine uptake and metabolism in marine microbial communities. _Environmental Microbiology_, 24(5), 2380-2403. [DOI:10.1111/1462-2920.16020](https://doi.org/10.1111/1462-2920.16020)
- **Kumler, W.E.**, Jorge, J., Kim, P.M., et al. (2020). Does formation of multicellular colonies by choanoflagellates affect their susceptibility to capture by passive protozoan predators? _Journal of Eukaryotic Microbiology_, 67(5), 555-565. [DOI:10.1111/jeu.12808](https://doi.org/10.1111/jeu.12808)


In review:

- E.A. Seelen, S.J. Gleich, **W. Kumler**, H.S. Anderson, X. Bian, K.M. Bjorkman, D.A. Caron, S.T. Dyhrman, S. Ferrón, Z.V. Finkel, S.T. Haley, Y. Hy, A.E. Ingalls, A.J. Irwin, D.M. Karl, K.P. Kong, D. Lowenstein, A. Salazar Estrada, E. Townsend, J.C. Tracey, K. Turk-Kubo, B.A.S. Van Mooy & S.G. John (in review with _Nature Communications_). A tale of two nutrients: how nitrogen and phosphorus differentially control marine biomass production and stoichiometry.

---

## Conference Proceedings

- **W. Kumler** & A.E. Ingalls. Databases are a speedy, small, and simple solution for mass-spectrometry data storage and access. Upcoming presentation at the _2025 Annual International Conference of the Metabolomics Society_, Prague, Czechia.
- P. Louail, **W. Kumler**, P. Vangeenderhuysen, C. Brunius, M. Witting, S. Neumann, RforMassSpectrometry contributors & J. Rainer. xcms at 20 and still in Peak form: Now anchoring a complete ecosystem for
metabolomics data preprocessing and analysis. Upcoming presentation at the _2025 Annual International Conference of the Metabolomics Society_, Prague, Czechia.
- **W. Kumler**, L.T. Carlson & A.E. Ingalls plus the PARAGON team. Differences in <sup>15</sup>N use by source during the PARAGON cruise in the NPSG. Presented as a poster at the _2024 Simons Collaboration on Ocean Processes and Ecology_, New York, NY.
- J. Rainer, P. Louail, A. Vicini, R. Gine, J. Badia, M. Stravs, M. Garcia-Aloy, C. Huber, L. Salzer, J. Stanstrup, N. Shahaf, C. Panse, T. Naake, **W. Kumler**, P. Vangeenderhuysen, C. Brunius, H. Hecht, S. Neumann, M. Witting, S. Gibb, & L. Gatto (2024). An Open Software Development-based Ecosystem of R Packages for Metabolomics Data Analysis. Presented as a poster at the _2024 Annual International Conference of the Metabolomics Society_, Osaka, Japan. DOI:10.5281/zenodo.13347220.
- S. Garcia, F.X. Ferrer-González, J.S. Sacks, **W. Kumler**, L.T. Carlson, & A.E. Ingalls (2024). Characterizing the quality and quantity of metabolite production in phytoplankton cultures and the environment through endo- and exometabolomics. Presented as a poster at the Gordon Research Conference on Marine Microbes, Les Diablerets, Switzerland. Poster #8.
- Y. Wang, **W. Kumler**, I. Kern, E. Seelen, S.G. John, & A.E. Ingalls (2023). Community metabolomes respond to nutrient supply in a mesocosm study in the North Pacific Subtropical Gyre. Presented as a poster at the _2023 Ocean Sciences Meeting_, New Orleans, LA. Poster #OB14B-0682.
- **W. Kumler**, L.T. Carlson, & A.E. Ingalls plus the PARAGON team (2023). Metabolic fate of dissolved nitrogen in the NPSG. Presented as a poster at the _2023 Ocean Sciences Meeting_, New Orleans, LA. Poster #OB14B-0702
- **W. Kumler**, L.T. Carlson, & A.E. Ingalls plus the PARAGON team (2023). Metabolic fate of dissolved nitrogen during PARAGON. Presented as a poster at the _2023 Simons Collaboration on Ocean Processes and Ecology_, New York, NY.
- **W. Kumler** (2022). Unmasking the POC/Ness Monster: Depth and Mesoscale Features Drive Variability in Particulate Metabolite Profiles of the MESOSCOPE transect. Presented as a virtual poster at the _2022 Simons Collaboration on Ocean Processes and Ecology_, New York, NY.
- **W. Kumler** (2022). Depth, diel, and eddy direction: The effect of three environmental factors on metabolite composition in the North Pacific Subtropical Gyre. Presented as a virtual poster at the _2021 Simons Collaboration on Ocean Processes and Ecology_, New York, NY.
- **W. Kumler** & A.E. Ingalls (2021). RaMS: R-based Access to Mass-Spectrometry Data. Presented as a poster at the _2021 Annual Conference of the Metabolomics Society (virtual)_. Poster #298.
- **W. Kumler**, H. Fredricks, J. Ossolinski, A. Allen, K. Thamatrakoln, K. Bidle, B. Van Mooy, & B. Edwards (2019). Sign of the times: the lipid signature of a collapsing phytoplankton bloom. Presented as a poster at the _2019 Aquatic Sciences Meeting_, San Juan, Puerto Rico. Poster #411.
- **W. Kumler** & M.A.R. Koehl (2018). Evolution of multicellularity: Capture of unicellular vs colonial choanoflagellates by a protozoan predator. Presented as a poster at the _2018 Society of Integrative and Comparative Biology Annual Meeting_, San Francisco, CA. Poster #P1-115.

---

## Talks Given

- **Building a robust model of peak quality for untargeted mass-spectrometry** (2023)  
    - Seminar given at UW's eScience Seminar (watch [here](https://www.youtube.com/watch?v=fNtp53o1wqQ&ab_channel=UWeScienceInstitute))
    - Seminar given at UW's Quantitative Science Seminar

- **Data Visualization** (2022)  
    - Workshop led at the Graduate Climate Conference

- **Profiling R code and identifying performance bottlenecks** (2019)  
    - Invited speaker at FSH 507 (Super-advanced R)

---

## Technical Projects

- **RaMS**: R package for reading and tidying mass-spectrometry data. Available on [CRAN](https://cran.r-project.org/package=RaMS) and [GitHub](https://github.com/wkumler/RaMS).
- **pylgrams**: Python language gloss of the RaMS package. Available on [PyPI](https://pypi.org/project/pylgrams/) and [GitHub](https://github.com/wkumler/pylgrams).
- **squallms**: R package for interactive multi-file annotation of chromatographic feature quality. Available on [Bioconductor](https://bioconductor.org/packages/devel/bioc/html/squallms.html) and [GitHub](https://github.com/wkumler/squallms).

---

## Awards

- Achievement Rewards for College Scientists ([ARCS](https://www.arcsfoundation.org/))
- [Mary C. Landsteiner Student Award](https://www.ocean.washington.edu/story/Mary_LandsteinerScholar)

---

## References

Available upon request.

